---
title: "GAGESiiExploring"
author: "Ben Choat"
date: "5/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Intro
This code is being written to work with and to explore GAGES ii data. 
Data exploration is focused around asking how the data can be used to predict 
daily, monthly, and annual water yield based on relevant characteristics in 
developed catchments and meteorological forcings using statistics and/or ML.

## Code

### Define functions and Load libraries
```{r}
if(!require(pacman))(install.packages("pacman"))
if(!require(devtools))(install.packages("devtools"))
if(!require(FedData))(devtools::install_github("ropensci/FedData"))

#data wrangling:
#data.table, tidyr, dplyr, dtplyr

#geospatial:
#sf: vectors and such
#terra: rasters and such
#USAboundaries and USAboundairesData: data and boundaires and such
#FedData: Data downloads

#Visualization and others:
#beepr: alarm when script is complete
#ggplot2: visualization
#tmap: geospatial visualization 
#readxl: read from excel .xls, .xlsx
pacman::p_load(data.table, tidyr, dplyr, dtplyr,
               sf, terra, USAboundaries, FedData,
               beepr, ggplot2, tmap, readxl)
```

### Define variables
```{r}
#working directory
wrk.dir <- "D:/"
```


### Load data
```{r}
#gages ii shapefile
ggs.shp <- sf::st_read(
  dsn = paste0(wrk.dir, "DataWorking/GAGESii/gagesII_9322_point_shapefile"), 
  layer = "gagesII_9322_sept30_2011")
#gages ii variable descriptions
ggs.var.disc <- data.table::data.table(readxl::read_excel(paste0(
  wrk.dir, "Projects/GAGESii_ANNstuff/gagesII_sept30_2011_var_desc_Exploring.xlsx"
    )
  )
)
#unique VARIABLE_TYPE gives tabs available in conterminous excel file
ggs.cont.tabs <- unique(ggs.var.disc$VARIABLE_TYPE)

#gages ii for conterminous U.S.
#filename; STAID is common key
ggs.cont.xcl <- paste0(
      wrk.dir, "DataWorking/GAGESii/basinchar_and_report_sept_2011/gagesII_sept30_2011_conterm.xlsx"
      )

#use variable descriptions and tabs to select what data to read in
#e.g.,
#ggs.var.disc[which(ggs.var.disc$VARIABLE_TYPE == "Bound_QA"), ][c(2,3,4)]

#Add the tabs you want to read in to the ggs.cont.tabs vector
ggs.cont.tabs <- c("BasinID", #lat, long, drainage area, HUC02, etc.
                   "Bas_Classif", #AGGECOREGION, HYDRO_DISTURB_INDX, comments, etc.
                   "Bas_Morph", #BAS_COPMACTNESS, LAT_CENT, LONG_CENT
                   "Bound_QA" #Confidence of basin boundary, drainage area
                   )
#Read in files as data.frames in a list, data.list
data.list <- lapply(ggs.cont.tabs, 
                    function(x){
                      readxl::read_excel(path = ggs.cont.xcl,
                     sheet = x)
                      }
                    )

#save as data.table
ggs.mstr <- data.table(Reduce(merge, data.list))
#subset to remove "poor" HUC10_CHECK"
ggs.rsnbl <- ggs.mstr[HUC10_CHECK %in% c("good", "reasonable")]
ggs.cnfdnc <- ggs.mstr$BASIN_BOUNDARY_CONFIDENCE
```

### Visualize
```{r}
#get state and city boundaries
st.brdrs.ggs <- USAboundaries::us_states(resolution = "low") %>% 
    filter(!state_abbr %in% c("HI", "PR", "AK"))

#plot gagesii stream gauge locations
ggs.map <- tmap::tm_shape(st.brdrs.ggs) +
  tm_polygons(col = "white") +
  tm_shape(ggs.shp) + 
  tm_dots(size = 0.01, col = "CLASS", palette = c("red", "blue"),
          legend.show = FALSE) +
  tm_add_legend(type = "symbol", title = "Class", 
                labels = c("Non-Ref (n = 7,265)", "Ref (n = 2,057)"), 
                col = c("red", "blue"), size = 1) +
  tm_layout(legend.text.size = 1, legend.position = c("left", "0.1")) +
  tm_compass(position = c("0.3", "bottom")) +
  tm_scale_bar(position = c("0", "bottom"))

ggs.map

# tmap::tmap_save(ggs.map, 
#                 filename = "E:/Projects/GAGESii_ANNstuff/outputs_visuals_andsuch/ggsii_cntgs.png")
  
```

### Some data exploration

```{r}
#unique variable types
unique(ggs.var.disc$VARIABLE_TYPE)

# return subset tables of variable descriptions
#var.types <- c("Climate_Ppt_Annual", "Climate_Tmp_Annual") #"Climate"
var.types <- c("Climate")
var.discr <- unique(
  ggs.var.disc[
    VARIABLE_TYPE %in% var.types,  
    .(VARIABLE_NAME, EXTENT, DESCRIPTION)
    ]
  )[
    order(rank(EXTENT))
    ]
```


