---
title: "GAGESii_Clustering_Wrk"
author: "Ben Choat"
date: "3/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Libraries and functions

```{r PackagesAndFunctions, echo=FALSE, warning=FALSE}
# use renv for R and R-package version control

#if(!require(renv))(install.pacakges("renv"))
#options for the renv package

# renv::init() #initiate new project
# renv::snapshot() #store active libraries in locked env. - Run this command anytime a package is updated or added.
# renv::restore() #restore libraries to previous stored state


# pacman as a package loader
if(!require(pacman))(install.packages("pacman"))

# renv for dependency management
# data.table for data wrangling
# ggplot for visualization
# ggiraph for interactive ggplots
# plotly for interactive plots
# corrplot for correlation and plots
# factoextra for PCA visualization
# Hmisc for caculating correlation matrix with p-values
# cluster for PAM and Gower space
# caret for splitting data into training, validating, and testing partitions
#   and other ML methods
pacman::p_load(renv, data.table, ggplot2, ggiraph, plotly, corrplot, factoextra,
               Hmisc, cluster, caret) 


```

### Read in data

```{r ReadData, warning=FALSE, echo=FALSE}
# annual water yield 
# training
dt_annWY_train <- data.table::fread(
        paste0(
        "D:/DataWorking/USGS_discharge/train_val_test/yrs_98_12/annual_WY/Ann_WY_train.csv"
        ) , keepLeadingZeros = TRUE
      )

# validate in
dt_annWY_valin <- data.table::fread(
        paste0(
        "D:/DataWorking/USGS_discharge/train_val_test/yrs_98_12/annual_WY/Ann_WY_val_in.csv"
        ) , keepLeadingZeros = TRUE
      )

# validate nit
dt_annWY_valnit <- data.table::fread(
        paste0(
        "D:/DataWorking/USGS_discharge/train_val_test/yrs_98_12/annual_WY/Ann_WY_val_nit.csv"
        ) , keepLeadingZeros = TRUE
      )

# static watershed characteristics
dt_static <- data.table::fread(
  "C:/R/GAGESii_R/Data_Out/GAGES_Static_Filtered.csv",
  keepLeadingZeros = TRUE
  )

# ID variables
dt_ID <- data.table::fread(
  "C:/R/GAGESii_R/Data_out/GAGES_idVars.csv",
  keepLeadingZeros = TRUE)

# calculate mean wy for each station
dt_annWYmn_train <- dt_annWY_train[
  , .(mn_anl_ft3 = mean(Ann_WY_ft3),
      mn_anl_acft = mean(Ann_WY_acft)),
      by = site_no]

dt_annWYmn_valin <- dt_annWY_valin[
  , .(mn_anl_ft3 = mean(Ann_WY_ft3),
      mn_anl_acft = mean(Ann_WY_acft)),
      by = site_no]

dt_annWYmn_train <- dt_annWY_train[
  , .(mn_anl_ft3 = mean(Ann_WY_ft3),
      mn_anl_acft = mean(Ann_WY_acft)),
      by = site_no]

dt_annWYmn_valnit <- dt_annWY_valnit[
  , .(mn_anl_ft3 = mean(Ann_WY_ft3),
      mn_anl_acft = mean(Ann_WY_acft)),
      by = site_no]


# data.table of static expl. vars and mean annual WY
dt_wy_static_train <- merge(
  dt_static,
  dt_annWYmn_train,
  by.x = "STAID", by.y = "site_no"
)

dt_wy_static_valin <- merge(
  dt_static,
  dt_annWYmn_valin,
  by.x = "STAID", by.y = "site_no"
)

dt_wy_static_valnit <- merge(
  dt_static,
  dt_annWYmn_valnit,
  by.x = "STAID", by.y = "site_no"
)

# long-format data.table of GAGESii-ts data with only years around those available
dt_srndg_train <- fread(
  "D:/Projects/GAGESii_ANNstuff/Data_Out/gagesii_ts/GAGESts_SurroundingYrs.csv",
  keepLeadingZeros = TRUE)[
   STAID %chin% dt_annWY_train$site_no &
    year %in% dt_annWY_train$wtryr , 
  ]

dt_srndg_valin <- fread(
  "D:/Projects/GAGESii_ANNstuff/Data_Out/gagesii_ts/GAGESts_SurroundingYrs.csv",
  keepLeadingZeros = TRUE)[
   STAID %chin% dt_annWY_valin$site_no &
    year %in% dt_annWY_valin$wtryr , 
  ]

dt_srndg_valnit <- fread(
  "D:/Projects/GAGESii_ANNstuff/Data_Out/gagesii_ts/GAGESts_SurroundingYrs.csv",
  keepLeadingZeros = TRUE)[
   STAID %chin% dt_annWY_valnit$site_no &
    year %in% dt_annWY_valnit$wtryr , 
  ]

# mean of explanatory vars

dt_ts_mn_train <- dt_srndg_train[
  , .(mean_value = mean(value)), by = c("STAID", "variable", "VOI")
  ]

dt_ts_mn_trainw <- data.table::dcast(
  dt_ts_mn_train, 
  STAID ~ VOI,
  value.var = "mean_value"
  )

dt_ts_mn_valin <- dt_srndg_valin[
  , .(mean_value = mean(value)), by = c("STAID", "variable", "VOI")
  ]

dt_ts_mn_valinw <- data.table::dcast(
  dt_ts_mn_valin, 
  STAID ~ VOI,
  value.var = "mean_value"
  )

dt_ts_mn_valnit <- dt_srndg_valnit[
  , .(mean_value = mean(value)), by = c("STAID", "variable", "VOI")
  ]

dt_ts_mn_valnitw <- data.table::dcast(
  dt_ts_mn_valnit, 
  STAID ~ VOI,
  value.var = "mean_value"
  )


# GAGESii-ts and static data and mean annual wy
dt_all_mn_train <- merge(
  dt_ts_mn_trainw,
  dt_wy_static_train,
  by = "STAID"
)[
  ,':=' (HYDRO_DISTURB_INDX = NULL,
      RUNAVE7100 = NULL)
]

dt_all_mn_valin <- merge(
  dt_ts_mn_valinw,
  dt_wy_static_valin,
  by = "STAID"
)[
  ,':=' (HYDRO_DISTURB_INDX = NULL,
      RUNAVE7100 = NULL)
]

dt_all_mn_valnit <- merge(
  dt_ts_mn_valnitw,
  dt_wy_static_valnit,
  by = "STAID"
)[
  ,':=' (HYDRO_DISTURB_INDX = NULL,
      RUNAVE7100 = NULL)
]

```


### Definw working data and investiagte how many stations are available at 
various levels of development.

### NOTE: stations 0164500 and 01646502 are identical based on gagesii static
variables and the averages of gages_ts variables over time. The mean annual
discharge over the training period is different by 0.06 ft.

```{r WorkData, echo=FALSE, warning=FALSE}
# which variables to remove
dt_wrk_train <- copy(dt_all_mn_train)[
  , ':=' (#TOPWET = NULL,
      GEOL_REEDBUSH_DOM = unclass(as.factor(GEOL_REEDBUSH_DOM)), #NULL,
      GEOL_HUNT_DOM_CODE = NULL,
      HYDRO_DISTURB_INDX = NULL,
      RUNAVE7100 = NULL,
      WB5100_ANN_MM = NULL,
      #STAID = NULL,
      CLASS = NULL,
      Class = NULL,
      BFI_AVE = NULL,
      #mn_anl_acft = NULL
      mn_anl_ft3 = NULL
      )
      ][
        , mn_anlQ_ft := mn_anl_acft/(DRAIN_SQKM * 247.105) 
      ][ # Note 274.105 acres in 1 km2 and 304.8 mm in 1 ft (* 304.8)
        , mn_anl_acft := NULL
      ]
dt_wrk_valin <- copy(dt_all_mn_valin)[
  , ':=' (#TOPWET = NULL,
      GEOL_REEDBUSH_DOM = unclass(as.factor(GEOL_REEDBUSH_DOM)), #NULL,
      GEOL_HUNT_DOM_CODE = NULL,
      HYDRO_DISTURB_INDX = NULL,
      RUNAVE7100 = NULL,
      WB5100_ANN_MM = NULL,
      #STAID = NULL,
      CLASS = NULL,
      Class = NULL,
      BFI_AVE = NULL,
      mn_anl_ft3 = NULL
      )
    ][
        , mn_anlQ_ft := mn_anl_acft/(DRAIN_SQKM * 247.105) 
      ][
        , mn_anl_acft := NULL
      ] 

dt_wrk_valnit <- copy(dt_all_mn_valnit)[
  , ':=' (#TOPWET = NULL,
      #GEOL_REEDBUSH_DOM = NULL,
      GEOL_HUNT_DOM_CODE = unclass(as.factor(GEOL_REEDBUSH_DOM)), #NULL,
      HYDRO_DISTURB_INDX = NULL,
      RUNAVE7100 = NULL,
      WB5100_ANN_MM = NULL,
      #STAID = NULL,
      CLASS = NULL,
      Class = NULL,
      BFI_AVE = NULL,
      mn_anl_ft3 = NULL
      )
    ][
        , mn_anlQ_ft := mn_anl_acft/(DRAIN_SQKM * 247.105) 
      ][
        , mn_anl_acft := NULL
      ] 
```


### Partitioning around  medoids in Gower space (aka PAM, k-medoids clustering)
leaning heavily on the websites:

1.
https://rpubs.com/sofiandra/477664
2.
https://dpmartin42.github.io/posts/r/cluster-mixed-types
3.
https://rpkgs.datanovia.com/factoextra/reference/fviz_silhouette.html

### paritioning around medoids with only GAGESii-static variables.

```{r PAM_static, echo=FALSE, warning=FALSE}
# data for clustering
data_in <- copy(dt_static)[
  STAID %chin% dt_all_mn_train$STAID,
][
  , ':=' (GEOL_REEDBUSH_DOM = as.factor(GEOL_REEDBUSH_DOM),
      GEOL_HUNT_DOM_CODE = NULL,
      HYDRO_DISTURB_INDX = NULL,
      STAID = NULL,
      CLASS = NULL,
      BFI_AVE = NULL,
      LAT_GAGE = NULL,
      LNG_GAGE = NULL
      #mn_anl_acft = NULL
      #mn_anl_ft3 = NULL
      )
]

# calculate gower distance
gower_dist <- cluster::daisy(data_in,
                             metric = "gower"
)

#gower_mat <- as.matrix(gower_dist)

# most similar based on distance calculation 
# data_in[
#   which(gower_mat == min(gower_mat[gower_mat != min(gower_mat)]),
#         arr.ind = TRUE)[1, ], ]
# 
# # most dissimilar based on distance calculation 
# data_in[
#   which(gower_mat == max(gower_mat[gower_mat != max(gower_mat)]),
#         arr.ind = TRUE)[1, ], ]


###
# Calculate silhouette width for many k using PAM

# sil_width <- c(NA)
# 
# for(i in 2:20){
# 
#   pam_fit <- cluster::pam(gower_dist,
#                  diss = TRUE,
#                  k = i)
# 
#   sil_width[i] <- pam_fit$silinfo$avg.width
# 
# }
# 
# # Plot sihouette width (higher is better)
# 
# plot(2:20, sil_width[2:20],
#      xlab = "Number of clusters",
#      ylab = "Silhouette Width")
# lines(2:20, sil_width[2:20])

## 
cat("Average silhouette width shows 2 (0.237) clusters to be the best, 
    followed by 3 (0.179), 9 (0.166), and 15 (0.169).
    All of these seem to be relatively low so, I explore 2, 3, 9, and 15")

## compute pam for specified 
for(i in c(2, 3, 9, 15)){
  
  tmp_pam <-  cluster::pam(gower_dist,
                 diss = TRUE,
                 k = i)
  
  assign(paste0("pam_fit_", i), tmp_pam)
  
  # assign(paste0("pam_results_", i), 
  #        data_in[
  #          , .(cluster = tmp_pam$clustering)
  #          ][
  #            , as.list(summary(data_in)), by = cluster
  #          ]
  #        )
  
  
  
}

factoextra::fviz_silhouette(pam_fit_2)
factoextra::fviz_silhouette(pam_fit_3)
factoextra::fviz_silhouette(pam_fit_9)
factoextra::fviz_silhouette(pam_fit_15)

pam_fit_2 <- cluster::pam(gower_dist,
                          diss = TRUE,
                          k = 2)
# look at general distribution of all vars
#data_in_long <- data.table::melt()
# 
# ggplot(data = data_in) +
#   geom_histogram(aes())

# data_in <- cor(copy(data_train_in)[, Qft := NULL])
# rownames(data_in) <- dt_ID[STAID %chin% dt_train$STAID, STAID]


# add cluster numbers to a data.table to keep record. Only keeping best
# performing K, in this case, 2

dt_clust_ID <- data.table::setorder(
  merge(
    data.table(
      STAID = copy(dt_static)[STAID %chin% dt_all_mn_train$STAID, STAID],
      PAM_2_Static_K = pam_fit_2$clustering
    ),
    data.table(
      PAM_2_Static_K = c(1, 2),
      PAM_2_Static_Med = pam_fit_2$medoids
    ), by = "PAM_2_Static_K"
  ),
  STAID
) # close setorder

# Write to a csv for later reference
data.table::fwrite(dt_clust_ID,
                "D:/Projects/GAGESii_ANNstuff/Data_Out/clustering_results.csv")

```


### paritioning around medoids with only GAGESii-ts variables.

```{r PAM_ts, echo=FALSE, warning=FALSE}
rm(list = ls(pattern = "pam_fit"))
set.seed(100)
# data for clustering
data_in <- copy(dt_ts_mn_trainw)[
  , STAID := NULL
]

# calculate gower distance
gower_dist <- cluster::daisy(data_in,
                             metric = "gower"
)

# gower_mat <- as.matrix(gower_dist)

# # most similar based on distance calculation 
# data_in[
#   which(gower_mat == min(gower_mat[gower_mat != min(gower_mat)]),
#         arr.ind = TRUE)[1, ], ]
# 
# # most dissimilar based on distance calculation 
# data_in[
#   which(gower_mat == max(gower_mat[gower_mat != max(gower_mat)]),
#         arr.ind = TRUE)[1, ], ]


###
# Calculate silhouette width for many k using PAM
# 
# sil_width <- c(NA)
# 
# for(i in 2:20){
# 
#   pam_fit <- cluster::pam(gower_dist,
#                  diss = TRUE,
#                  k = i)
# 
#   sil_width[i] <- pam_fit$silinfo$avg.width
# 
# }

# Plot sihouette width (higher is better)
# 
# plot(2:20, sil_width[2:20],
#      xlab = "Number of clusters",
#      ylab = "Silhouette Width")
# lines(2:20, sil_width[2:20])
# 
# ## 
# cat("Average silhouette width shows 2 (0.432) clusters to be the best, 
#     followed by 4 (0.38).")

## compute pam for specified 
for(i in c(2, 4)){
  
  tmp_pam <-  cluster::pam(gower_dist,
                 diss = TRUE,
                 k = i)
  
  assign(paste0("pam_fit_", i), tmp_pam)
  
  # assign(paste0("pam_results_", i), 
  #        data_in[
  #          , .(cluster = pam_fit$clustering)
  #          ][
  #            , as.list(summary(data_in)), by = cluster
  #          ]
  #        )
  
  
  
}


factoextra::fviz_silhouette(pam_fit_2)
factoextra::fviz_silhouette(pam_fit_4)

pam_fit_2 <- cluster::pam(gower_dist,
                          diss= TRUE,
                          k = 2)

# add cluster numbers to a data.table to keep record. Only keeping best
# performing K

dt_clust_ID <- data.table::setorder(
  merge(
    data.table(
      STAID = dt_wy_static_train$STAID,
      PAM_2_TS_K = pam_fit_2$clustering
    ),
    data.table(
      PAM_2_TS_K = c(1, 2),
      PAM_2_TS_Med = pam_fit_2$medoids
    ), by = "PAM_2_TS_K"
  ),
  STAID
) # close setorder

tmp_clust_rslts <- data.table::fread(
  "D:/Projects/GAGESii_ANNstuff/Data_Out/clustering_results.csv",
  keepLeadingZeros = TRUE)

tmp_dt_write <- merge(
  tmp_clust_rslts,
  dt_clust_ID,
  by = "STAID"
)

# Write to a csv for later reference
data.table::fwrite(tmp_dt_write,
                "D:/Projects/GAGESii_ANNstuff/Data_Out/clustering_results.csv")

# look at general distribution of all vars
#data_in_long <- data.table::melt()
# 
# ggplot(data = data_in) +
#   geom_histogram(aes())

# data_in <- cor(copy(data_train_in)[, Qft := NULL])
# rownames(data_in) <- dt_ID[STAID %chin% dt_train$STAID, STAID]

```

### paritioning around medoids with all GAGESii-ts variables.

```{r PAM_all, echo=FALSE, warning=FALSE}
rm(list = ls(pattern = "pam_fit"))
set.seed(100)
# data for clustering
data_in <- copy(dt_wrk_train)[
  , ':=' (STAID = NULL,
           mn_anlQ_ft = NULL)
]

# calculate gower distance
gower_dist <- cluster::daisy(data_in,
                             metric = "gower"
)

# gower_mat <- as.matrix(gower_dist)

# most similar based on distance calculation 
# data_in[
#   which(gower_mat == min(gower_mat[gower_mat != min(gower_mat)]),
#         arr.ind = TRUE)[1, ], ]
# 
# # most dissimilar based on distance calculation 
# data_in[
#   which(gower_mat == max(gower_mat[gower_mat != max(gower_mat)]),
#         arr.ind = TRUE)[1, ], ]


###
# Calculate silhouette width for many k using PAM
silh_plot <- factoextra::fviz_nbclust(gower_mat, pam, method='silhouette', k.max = 2)
silh_plot

pam_fit_2 <- cluster::pam(gower_dist,
                          diss = TRUE,
                          k = 2)

factoextra::fviz_silhouette(pam_fit_2)
factoextra::fviz_silhouette(pam_fit_4)

# add cluster numbers to a data.table to keep record. Only keeping best
# performing K

dt_clust_ID <- data.table::setorder(
  merge(
    data.table(
      STAID = dt_wy_static_train$STAID,
      PAM_2_All_K = pam_fit_2$clustering
    ),
    data.table(
      PAM_2_All_K = c(1, 2),
      PAM_2_All_Med = pam_fit_2$medoids
    ), by = "PAM_2_All_K"
  ),
  STAID
) # close setorder

tmp_clust_rslts <- data.table::fread(
  "D:/Projects/GAGESii_ANNstuff/Data_Out/clustering_results.csv",
  keepLeadingZeros = TRUE)

tmp_dt_write <- merge(
  tmp_clust_rslts,
  dt_clust_ID,
  by = "STAID"
)


# Write to a csv for later reference
data.table::fwrite(tmp_dt_write,
                "D:/Projects/GAGESii_ANNstuff/Data_Out/clustering_results.csv")

# look at general distribution of all vars
#data_in_long <- data.table::melt()
# 
# ggplot(data = data_in) +
#   geom_histogram(aes())

# data_in <- cor(copy(data_train_in)[, Qft := NULL])
# rownames(data_in) <- dt_ID[STAID %chin% dt_train$STAID, STAID]

```

### Cluster the original clusters to see how they perform

```{r CulsterClustersStatic, warning=FALSE, echo=FALSE}
dt_clsts <- data.table::fread(
  "D:/Projects/GAGESii_ANNstuff/Data_Out/clustering_results.csv",
  keepLeadingZeros = TRUE
)

data_in <- copy(dt_static)[
  STAID %chin% dt_all_mn_train$STAID,
][
  , ':=' (GEOL_REEDBUSH_DOM = as.factor(GEOL_REEDBUSH_DOM),
      GEOL_HUNT_DOM_CODE = NULL,
      HYDRO_DISTURB_INDX = NULL,
      CLASS = NULL,
      BFI_AVE = NULL,
      LAT_GAGE = NULL,
      LNG_GAGE = NULL
      #mn_anl_acft = NULL
      #mn_anl_ft3 = NULL
      )
]

data_in <- merge(
  data_in,
  dt_clsts[, .(STAID, PAM_2_Static_K)],
  by = "STAID"
)

data_in_k1 <- copy(data_in)[
  PAM_2_Static_K == 1,
][
  , ':=' (STAID = NULL,
          PAM_2_Static_K = NULL)
]


data_in_k2 <- copy(data_in)[
  PAM_2_Static_K == 2,
][
  , ':=' (STAID = NULL,
          PAM_2_Static_K = NULL)
]


# calculate gower distance
gower_dist_k1 <- cluster::daisy(data_in_k1,
                             metric = "gower",
                             stand = TRUE
)

gower_dist_k2 <- cluster::daisy(data_in_k2,
                             metric = "gower",
                             stand = TRUE
)

# visualize distances between variables
DistanceMap <- fviz_dist(
  gower_dist_k1, 
  order = TRUE, 
  show_labels = TRUE, 
  lab_size = 4) + 
  labs(title = "Distance: Gower Measurements")

###
# Calculate silhouette width for many k using PAM
# 
sil_width <- c(NA)

for(i in 2:10){

  pam_fit <- cluster::pam(gower_dist_k2,
                diss = TRUE,
                 k = i,
                 )

  sil_width[i] <- pam_fit$silinfo$avg.width

}

# Plot sihouette width (higher is better)

plot(2:10, sil_width[2:10],
     xlab = "Number of clusters",
     ylab = "Silhouette Width")
lines(2:10, sil_width[2:10])

# ## 
# cat("Average silhouette width shows 2 (0.432) clusters to be the best, 
#     followed by 4 (0.38).")

## compute pam for specified 
for(i in c(2, 4)){
  
  tmp_pam <-  cluster::pam(gower_dist,
                 diss = TRUE,
                 k = i)
  
  assign(paste0("pam_fit_", i), tmp_pam)

}

# nb_k1 <- factoextra::fviz_nbclust()

silh_plot <- factoextra::fviz_nbclust(
  x = as.matrix(gower_dist_k2), #data_in_k2,
  #diss = as.matrix(gower_dist_k2), 
  FUNcluster = pam, 
  method = 'gap_stat',
  k.max = 10
  )
```











<!-- ### k-means clustering with only GAGESii-static variables. -->

<!-- ```{r kmeans_static, echo=FALSE, warning=FALSE} -->
<!-- # data for clustering -->
<!-- data_in <- copy(dt_static)[ -->
<!--   STAID %chin% dt_all_mn_train$STAID, -->
<!-- ][ -->
<!--   , ':=' (GEOL_REEDBUSH_DOM = as.factor(GEOL_REEDBUSH_DOM), -->
<!--       GEOL_HUNT_DOM_CODE = NULL, -->
<!--       HYDRO_DISTURB_INDX = NULL, -->
<!--       STAID = NULL, -->
<!--       CLASS = NULL, -->
<!--       BFI_AVE = NULL, -->
<!--       LAT_GAGE = NULL, -->
<!--       LNG_GAGE = NULL -->
<!--       #mn_anl_acft = NULL -->
<!--       #mn_anl_ft3 = NULL -->
<!--       ) -->
<!-- ] -->
<!-- # calculate gower distance -->
<!-- gower_dist <- cluster::daisy(data_in, -->
<!--                              metric = "gower" -->
<!-- ) -->

<!-- gower_mat <- as.matrix(gower_dist) -->


<!-- ### -->
<!-- # Calculate silhouette width for many k using kmeans -->

<!-- silh_plot <- factoextra::fviz_nbclust(gower_mat, kmeans, method='silhouette', k.max = 10) -->

<!-- km_fit_2 <- kmeans(x = gower_mat, -->
<!--                    centers = 2) -->

<!-- # add cluster numbers to a data.table to keep record. Only keeping best -->
<!-- # performing K -->

<!-- dt_clust_ID <- data.table::setorder( -->
<!--   merge( -->
<!--     data.table( -->
<!--       STAID = dt_wy_static_train$STAID, -->
<!--       KM_2_Static_K = km_fit_2$cluster -->
<!--     ), -->
<!--     data.table( -->
<!--       KM_2_Static_K = c(1, 2), -->
<!--       KM_2_Static_Mn = km_fit_2$medoids -->
<!--     ), by = "KM_2_Static_K" -->
<!--   ), -->
<!--   STAID -->
<!-- ) # close setorder -->

<!-- tmp_clust_rslts <- data.table::fread( -->
<!--   "D:/Projects/GAGESii_ANNstuff/Data_Out/clustering_results.csv", -->
<!--   keepLeadingZeros = TRUE) -->

<!-- tmp_dt_write <- merge( -->
<!--   tmp_clust_rslts, -->
<!--   dt_clust_ID, -->
<!--   by = "STAID" -->
<!-- ) -->

<!-- # Write to a csv for later reference -->
<!-- data.table::fwrite(tmp_dt_write, -->
<!--                 "D:/Projects/GAGESii_ANNstuff/Data_Out/clustering_results.csv") -->

<!-- ``` -->

<!-- ### k-means clustering with only GAGESii-TS variables. -->

<!-- ```{r kmeans_TS, echo=FALSE, warning=FALSE} -->
<!-- # data for clustering -->
<!-- rm(list = ls(pattern = "km_fit")) -->
<!-- set.seed(100) -->
<!-- # data for clustering -->
<!-- data_in <- copy(dt_ts_mn_trainw)[ -->
<!--   , STAID := NULL -->
<!-- ] -->

<!-- # calculate gower distance -->
<!-- gower_dist <- cluster::daisy(data_in, -->
<!--                              metric = "gower" -->
<!-- ) -->

<!-- gower_mat <- as.matrix(gower_dist) -->


<!-- ### -->
<!-- # Calculate silhouette width for many k using kmeans -->

<!-- silh_plot <- factoextra::fviz_nbclust(gower_mat, -->
<!--                                       kmeans,  -->
<!--                                       method='silhouette',  -->
<!--                                       k.max = 10) -->


<!-- km_fit_2 <- kmeans(x = gower_mat, -->
<!--                    centers = 2) -->

<!-- km_fit_3 <- kmeans(x = gower_mat, -->
<!--                    centers = 3) -->

<!-- sil <- silhouette(km_fit_2$cluster, gower_dist) -->
<!-- factoextra::fviz_silhouette(sil) -->

<!-- sil <- silhouette(km_fit_3$cluster, gower_dist) -->
<!-- factoextra::fviz_silhouette(sil) -->

<!-- # add cluster numbers to a data.table to keep record. Only keeping best -->
<!-- # performing K -->

<!-- dt_clust_ID <- data.table::setorder( -->
<!--   merge( -->
<!--     data.table( -->
<!--       STAID = dt_wy_TS_train$STAID, -->
<!--       KM_2_TS_K = km_fit_2$cluster -->
<!--     ), -->
<!--     data.table( -->
<!--       KM_2_TS_K = c(1, 2), -->
<!--       KM_2_TS_Mn = pam_fit_2$medoids -->
<!--     ), by = "KM_2_TS_K" -->
<!--   ), -->
<!--   STAID -->
<!-- ) # close setorder -->

<!-- tmp_clust_rslts <- data.table::fread( -->
<!--   "D:/Projects/GAGESii_ANNstuff/Data_Out/clustering_results.csv", -->
<!--   keepLeadingZeros = TRUE) -->

<!-- tmp_dt_write <- merge( -->
<!--   tmp_clust_rslts, -->
<!--   dt_clust_ID, -->
<!--   by = "STAID" -->
<!-- ) -->

<!-- # Write to a csv for later reference -->
<!-- data.table::fwrite(tmp_dt_write, -->
<!--                 "D:/Projects/GAGESii_ANNstuff/Data_Out/clustering_results.csv") -->

<!-- ``` -->


<!-- ### Hierarchical clustering on GAGES-static variables -->

<!-- ```{r HiClust_static, echo=FALSE, warning=FALSE} -->

<!-- data_in <- copy(dt_static_train)[ -->
<!--   , ':=' (GEOL_REEDBUSH_DOM = NULL, #as.factor(GEOL_REEDBUSH_DOM), -->
<!--       GEOL_HUNT_DOM_CODE = NULL, -->
<!--       HYDRO_DISTURB_INDX = NULL, -->
<!--       STAID = NULL, -->
<!--       CLASS = NULL, -->
<!--       BFI_AVE = NULL, -->
<!--       LAT_GAGE = NULL, -->
<!--       LNG_GAGE = NULL -->
<!--       #mn_anl_acft = NULL -->
<!--       #mn_anl_ft3 = NULL -->
<!--       ) -->
<!-- ] -->

<!-- rownames(data_in) <- dt_static_train[STAID %chin% dt_ID_train$STAID, STAID] -->

<!-- # scale the data then calculate the distance matrix -->
<!-- dist_in <- dist(scale(data_in), method = "euclidian") -->
<!-- #dist_in <- cluster::daisy(data_in, metric = "gower", stand = TRUE) -->
<!-- #dist_in <- as.dist(1 - cor(data_in)) #, method = "euclidian") -->

<!-- # hierarchical cluster -->
<!-- out_hclust <- hclust(dist_in, method = "complete") -->


<!-- plot(out_hclust, cex = 0.02)#, hang = -1) -->
<!-- ``` -->

<!-- ### Hierarchical clustering on GAGES-TS variables -->

<!-- ```{r HiClust_ts, echo=FALSE, warning=FALSE} -->

<!-- data_in <- copy(dt_ts_mn_train)[ -->
<!--   , STAID := NULL -->
<!-- ] -->

<!-- rownames(data_in) <- dt_static_train[STAID %chin% dt_ID_train$STAID, STAID] -->

<!-- # scale the data then calculate the distance matrix -->
<!-- dist_in <- dist(scale(data_in), method = "euclidian") -->
<!-- #dist_in <- cluster::daisy(data_in, metric = "gower", stand = TRUE) -->
<!-- #dist_in <- as.dist(1 - cor(data_in)) #, method = "euclidian") -->

<!-- # hierarchical cluster -->
<!-- out_hclust <- hclust(dist_in, method = "complete") -->


<!-- plot(out_hclust, cex = 0.02)#, hang = -1) -->
<!-- ``` -->


### The Code
```{r ref.label="PackagesAndFunctions", eval=FALSE}
```

```{r ref.label="ReadData", eval=FALSE}
```

```{r ref.label="WorkData", eval=FALSE}
```

```{r ref.label="PAM_static", eval=FALSE}
```

```{r ref.label="PAM_ts", eval=FALSE}
```

```{r ref.label="PAM_all", eval=FALSE}
```

```{r ref.label="HiClust_static", eval=FALSE}
```

```{r ref.label="HiClust_ts", eval=FALSE}
```