---
title: "GatherStreamflow"
author: "Ben Choat"
date: "2/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### A short script to download usgs daily streamflow data using the USGS 
### dataretrieval package.

### Load packages

```{r PackagesAndFunctions, echo=FALSE, warning=FALSE}
# use renv for R and R-package version control

#if(!require(renv))(install.pacakges("renv"))
#options for the renv package

# renv::init() #initiate new project
# renv::snapshot() #store active libraries in locked env. - Run this command anytime a package is updated or added.
# renv::restore() #restore libraries to previous stored state


# pacman as a package loader
if(!require(pacman))(install.packages("pacman"))

# renv for dependency management
# data.table for data wrangling
# dataRetrieval as USGS data API
# ggplot2 for plotting
# plotly for plotting


pacman::p_load(renv, data.table, dataRetrieval, ggplot2, plotly)

```

### Read in the gages id variables to identify stations to download
```{r readData, echo=FALSE, warning=FALSE}
dt_STAID <- data.table::fread(
  "D:/Projects/GAGESii_ANNstuff/Data_Out/GAGES_Cont_STATID.csv",
  keepLeadingZeros = TRUE
  )

dt_static <- data.table::fread(
  "C:/R/GAGESii_R/Data_Out/GAGES_Static_Filtered.csv",
  keepLeadingZeros = TRUE
  )[
    STAID %chin% dt_STAID$STAID
  ]

dt_ID <- data.table::fread(
  "C:/R/GAGESii_R/Data_out/GAGES_idVars.csv",
  keepLeadingZeros = TRUE)[
    STAID %chin% dt_STAID$STAID,
    ]


```


### Download streamflow data at available resolution. 
### Caution: If you run this script 1000's of csv's will be produced and 
### the script could take hours.

I attempted to download sub-daily data because USGS streamflow data
and DAYMET data may not be (are not) aligned to the same time zone. But, 
sub-daily data is not available for the periods indicated in GAGESii - only mean
daily discharge is, so that is what I downloaded.

Data will be downloaded to the DataRaw folder on my external hard drive, in case
it is of interest in future work.

A csv will be written to hold a list of stations downloaded.

```{r StreamflowDownload, echo=FALSE, warning=FALSE}
system.time({
# Define input vars specifying what and from when to download data
siteNo <- dt_ID$STAID[1]
pCode <- "00060" # instantaneous discharge ft3/s
start_date <- "1976-01-01"
end_date <- "2013-12-31"


for(siteNo in dt_ID$STAID){#seq(1, 3, 1)) { # 203
  
  tmp_dschrg <- as.data.table(
    readNWISdv(
      siteNumbers = siteNo,
      parameterCd = pCode,
      startDate = start_date,
      endDate = end_date
    )
  )
  
  data.table::fwrite(tmp_dschrg,
    paste0(
      "D:/DataRaw/USGS_discharge/daily_mean/Daily_Mean_Q_1976_2013_",
      siteNo, 
      ".csv"
      )
  )
  data.table::fwrite(tmp_dschrg,
    paste0(
      "D:/DataWorking/USGS_discharge/daily_mean/Daily_Mean_Q_1976_2013_",
      siteNo, 
      ".csv"
      )
  )

}

})

# see what units are
tmp_info <- readNWISpCode(parameterCd = "00060")
```


### Produce summary csv with data from the dt_ID data.table as well as start and
### end dates. 

I probably should have done this during the original download, but I didn't.

1. Loop through station ID's.
2. For each continuous group of dates, add a row with the station ID, start
  date(s), and end date(s)
3. If the CSV is empty, add the NA

```{r MarkDates, echo=FALSE, warning=FALSE}
system.time({
# initiate empty data.table to add  data to
tmp_dts_run <- data.table(
  STAID = character(), 
  Start_Dates = character(), 
  End_Dates = character(), 
  Consec_Days = numeric()
  )

# Loop through stations and identify start and end dates of continuous data.
for(siteNo in dt_ID$STAID){ # c("01022260", "07307010")
  tmp <- data.table::fread(
          paste0(
            "D:/DataWorking/USGS_discharge/daily_mean/Daily_Mean_Q_1976_2013_",
            siteNo, 
            ".csv"
          ),
          keepLeadingZeros = TRUE
        )
  
  tmp$Date <- as.Date(tmp$Date)
  # create vector notating which dates are consecutive
  test <- c(NA, diff(tmp$Date) == 1)
  
  # create indices of first and last dates of continuous dates
  tmp_indx <- which(!test == TRUE)
  tmp_indx <- c(tmp_indx, tmp_indx - 1)
  tmp_indx <- tmp_indx[which(tmp_indx > 0)]
  tmp_indx <- sort(c(tmp_indx, 1, length(test)))
  
  if(all(tmp_indx == 1)) {
    tmp_dts <- data.table(
      STAID = siteNo,
      Start_Dates = NA,
      End_Dates = NA,
      Consec_Days = 0
    )
  } else {
    tmp_start_dates <- tmp$Date[tmp_indx[seq(1, length(tmp_indx), 2)]]
    tmp_end_dates <- tmp$Date[tmp_indx[seq(2, length(tmp_indx), 2)]]
    tmp_consec_days <- tmp_end_dates - tmp_start_dates + 1
    
    tmp_dts <- data.table(
      STAID = rep(siteNo, length(tmp_indx)/2),
      Start_Dates = as.character(tmp_start_dates),
      End_Dates = as.character(tmp_end_dates))[
      , Consec_Days := as.numeric(tmp_consec_days)
      ]
    
  }
  
  
  # # write to csv
  # data.table::fwrite(
  #   tmp_dts,
  #   "D:/Projects/GAGESii_ANNstuff/Data_Out/Daily_Q_Dates.csv",
  #   append = TRUE
  #   )
  
  tmp_dts_run <- rbind(tmp_dts_run, tmp_dts)
}

  # write to csv
  data.table::fwrite(
    tmp_dts_run,
    "D:/Projects/GAGESii_ANNstuff/Data_Out/Daily_Q_Dates.csv",
    append = TRUE
    )
  
})

```
### Calculate total daily, monthly, and annual water yield
Exclude Dec. 31 on leap years to match DAYMET weather data.
Write daily WY data to individual files for each station, but
write all monthly and annual wy data to single tables - one for monthly and one
for annual WY.

Set up to keep only discharge where there are at least three continuous
water years of discharge.

```{r SumWY, echo=FALSE, warning=FALSE}
system.time({
# define list specifying number of days in each month
dy30 <- c(4, 6, 9, 11)
dy31 <- c(1, 3, 5, 7, 8, 10, 12)

# # read in start and end dates and add columns for start and end days, months,
# # and years
# dt_dts <- fread("D:/Projects/GAGESii_ANNstuff/Data_Out/Daily_Q_Dates.csv",
#                keepLeadingZeros = TRUE)[
#                  , ':=' (
#                    st_day = substr(Start_Dates, 9, 10),
#                    end_day = substr(End_Dates, 9, 10),
#                    st_mnth = substr(Start_Dates, 6, 7),
#                    end_mnth = substr(End_Dates, 6, 7),
#                    st_yr = substr(Start_Dates, 1, 4),
#                    end_yr = substr(End_Dates, 1, 4)
#                    )
#                ]
  
# initiate empty data.table to add monthly wy data to
tmp_mnthwy_run <- data.table(
  site_no = character(),
  yr = character(),
  wtryr = character(),
  mnth = character(),
  Mnth_WY_ft3 = numeric(),
  Mnth_WY_acft = numeric()
  )
# initiate empty data.table to add annual wy data to
tmp_annwy_run <- data.table(
  site_no = character(),
  yr = character(),
  Ann_WY_ft3 = numeric(),
  Ann_WY_acft = numeric()
  )

# initiate vector to write empty stations to
st_empty <- c()


# Loop through stations and identify start and end dates of continuous data.
for(siteNo in dt_ID$STAID){ #"13214000"){# c("01022260", "07307010")
  
  # read in mean discharge for siteNo
  tmp <- data.table::fread(
          paste0(
            "D:/DataWorking/USGS_discharge/daily_mean/Daily_Mean_Q_1976_2013_",
            siteNo, 
            ".csv"
          ),
          keepLeadingZeros = TRUE
        )
  
  # if tmp is empty, then next siteNo
  if(nrow(tmp) == 0){
    st_empty <- c(st_empty, siteNo)
    next
  }
  
  # make sure siteno is a character (at least for one gauge it was saved as int64)
  tmp$site_no <- as.character(tmp$site_no)
  # Add columns for year and month individually
  tmp[
    , ':=' (
      yr = substr(tmp$Date, 1, 4),
      mnth = substr(tmp$Date, 6, 7),
      day = substr(tmp$Date, 9, 10)
    )][
      , ':=' (wtryr = ifelse(
        as.numeric(mnth) %between% c(9, 12),
        as.character(as.numeric(yr) + 1),
        yr
        )
      )
    ]
  
  # which months to keep? - Only months with an expected number of days
  mnth_kp <- tmp[
    , length(unique(day)), 
    by = .(yr, mnth)
    ][
      mapply(function(x, y) {
        if(as.numeric(y) %in% dy30){
          x == 30
        } else if (as.numeric(y) %in% dy31){       
          x == 31
        } else {
          x %in% c(28, 29)
        }
          },
          V1, mnth   
             ),
    ]
  
  # which years to keep? - Those with all dates still present
  yr_kp <- data.table(
    yr = unique(
      tmp$yr[
        tmp$yr %chin% mnth_kp[, yr[length(mnth) == 12],
                                 by =yr]$V1
      ]
    ) # close unique
  )
  # Commented out selection here, was used when I previously filtered by 
  # water year instead of julian year.
  # [# close data.table
  #   ,.(yr,
  #      crtr1 = as.numeric(yr) - as.numeric(shift(yr, 2, fill = min(as.numeric(yr), na.rm = TRUE))),
  #      crtr2 = as.numeric(yr) - as.numeric(shift(yr, -2, fill = max(as.numeric(yr), na.rm = TRUE)))
  #      )
  # ][
  #   !crtr1 == 0 & !crtr2 == 0 & (abs(crtr1) <= 2 | abs(crtr2 <= 2)) | 
  #     crtr1 == 0 & abs(crtr2) <= 2 | 
  #     crtr2 == 0 & abs(crtr1) <= 2,
  # ]

  
  

  tmp <- tmp[
    which(mapply(
      function(x, y) {
        x %chin% yr_kp$yr &
        y %chin% mnth_kp$mnth
      }, 
     yr, mnth
    )),
    ]
  
  # calculate daily wy and add to tmp data.table
  if("X_..2.._00060_00003" %chin% colnames(tmp)){
      tmp$dlyWY_cfd <- tmp$X_..2.._00060_00003 * 86400
  }else if("X_00060_00003" %chin% colnames(tmp)){
    tmp$dlyWY_cfd <- tmp$X_00060_00003 * 86400
  }else if("X_PUBLISHED_00060_00003" %chin% colnames(tmp)){
      tmp$dlyWY_cfd <- tmp$X_PUBLISHED_00060_00003 * 86400
  }else if("X_.Primary.Stream.Flow._00060_00003" %chin% colnames(tmp)){
      tmp$dlyWY_cfd <- tmp$X_.Primary.Stream.Flow._00060_00003 * 86400
  }
  

  # subset tmp to complete years
  #tmp_mnth_cmpl <- tmp[!]
  
  # calcualte monthly wy (43560 ft2 in acre)
  tmp_mnthwy_run <- rbind(
    tmp_mnthwy_run,
    tmp[
    , .(Mnth_WY_ft3 = sum(dlyWY_cfd),
        Mnth_WY_acft = sum(dlyWY_cfd)/43560,
        wtryr = ifelse(
          as.numeric(mnth) %between% c(9, 12),
          as.character(as.numeric(yr) + 1),
          yr
        )),
    by = .(site_no, yr,mnth)
    ]
  )
  # calculate annual wy
  tmp_annwy_run <- rbind(
    tmp_annwy_run,
    tmp[
    , .(Ann_WY_ft3 = sum(dlyWY_cfd),
        Ann_WY_acft = sum(dlyWY_cfd)/43560
        ),
    by = .(site_no, yr)
    ]
  )
  
  data.table::fwrite(tmp,
    paste0(
      "D:/DataWorking/USGS_discharge/daily_WY/Daily_WY_1976_2013_",
      siteNo,
      ".csv"
      )
    )
}
  
   data.table::fwrite(tmp_mnthwy_run,
    paste0(
      "D:/DataWorking/USGS_discharge/monthly_WY/Monthly_WY_1976_2013.csv"
      )
    )
   
   data.table::fwrite(tmp_annwy_run,
    paste0(
      "D:/DataWorking/USGS_discharge/annual_WY/Annual_WY_1976_2013.csv"
      )
    )

  
}) # system.time
```


### Filter/explore data

```{r Explr, echo=FALSE, warning=FALSE}
dt_dts <- fread("D:/Projects/GAGESii_ANNstuff/Data_Out/Daily_Q_Dates.csv",
               keepLeadingZeros = TRUE)
#dt_dts <- fread("C:/R/GAGESii_R/Data_Out/Daily_Q_Dates.csv",
#                keepLeadingZeros = TRUE)

# timespans for each station less than 4 years
dt_dts_del <- dt_dts[Consec_Days < (4 * 365)]
# timespans for each station greater than 4 years
dt_dts_keep <- dt_dts[Consec_Days >= (4 * 365)]

nmbr_sttns_del <- length(unique(dt_dts_del$STAID))
nmbr_sttns_kp <- length(unique(dt_dts_keep$STAID))
```

### Read in mean-daily discharge data and create some plots

```{r PlotQmean, warning=FALSE, echo=FALSE}
siteNo <- dt_ID$STAID[1:5]

# read in discharge data from specified stations to a single data.table
dt_Q <- data.table::rbindlist(
  lapply(
    siteNo,
    function(x) {
      data.table::fread(
        paste0(
        "D:/DataWorking/USGS_discharge/daily_mean/Daily_Mean_Q_1976_2013_",
        x, ".csv"
        # "C:/R/GAGESii_R/USGS_discharge/daily_mean/Daily_Mean_Q_1976_2013_",
        # x, ".csv"
        ) #, keepLeadingZeros = TRUE
        )
    }
  )
)

# dt_Qmap <- dt_Q
# dt_Qmap$site_no <- as.numeric(dt_Qmap$site_no)

ggplot(data = dt_Q) +
  geom_line(aes(x = Date, 
                y = X_00060_00003,
                linetype = as.factor(site_no),
                color = as.factor(site_no))) +
  ylab("Mean Daily \n Discharge [ft3/s]") +
  theme_bw()
```


### Read in daily WY data and create some plots
```{r plotdailyWY, warning=FALSE, echo=FALSE}
siteNo <- dt_ID$STAID[1:10]

# read in discharge data from specified stations to a single data.table
dt_dlyWY <- data.table::rbindlist(
  lapply(
    siteNo,
    function(x) {
      data.table::fread(
        paste0(
        "D:/DataWorking/USGS_discharge/daily_WY/Daily_WY_1976_2013_",
        x, ".csv"
        # "C:/R/GAGESii_R/USGS_discharge/daily_mean/Daily_Mean_Q_1976_2013_",
        # x, ".csv"
        ) , keepLeadingZeros = TRUE
        )
    }
  )
)

ggplot(data = dt_dlyWY) +
  geom_line(aes(x = as.Date(Date), 
                y = dlyWY_cfd/43560,
                linetype = as.factor(site_no),
                color = as.factor(site_no))) +
  ylab("Daily WY [acre-ft]") +
  theme_bw()



```

### Read in monthly WY data and create some plots
```{r plotMonthlyWY, warning=FALSE, echo=FALSE}
# read in monthly wy
dt_mnthlyWY <- data.table::fread(
        paste0(
        "D:/DataWorking/USGS_discharge/monthly_WY/Monthly_WY_1976_2013.csv"
        ) , keepLeadingZeros = TRUE
        )
    
dt_mnthlyWY$Mnth_Yr <- as.Date(
  paste0(dt_mnthlyWY$wtryr, "-", dt_mnthlyWY$mnth, "-", "01"),
  format = "%Y-%m-%d"
  )

siteNo <- dt_ID$STAID[1:10]


p <- ggplot(data = dt_mnthlyWY[site_no %in% siteNo, ]) +
  geom_line(aes(x = Mnth_Yr, 
                y = Mnth_WY_acft,
                linetype = as.factor(site_no),
                color = as.factor(site_no))) +
  ylab("Monthly WY [acre-ft]") +
  theme_bw()


plotly::ggplotly(p)
```

### Read in annual WY data and create some plots
```{r plotMonthlyWY, warning=FALSE, echo=FALSE}
# read in monthly wy
dt_annWY <- data.table::fread(
        paste0(
        "D:/DataWorking/USGS_discharge/annual_WY/Annual_WY_1976_2013.csv"
        ) , keepLeadingZeros = TRUE
        )
    
siteNo <- dt_ID$STAID[1:10]


p <- ggplot(data = dt_annWY[site_no %in% siteNo, ]) +
  geom_line(aes(x = wtryr, 
                y = Ann_WY_acft,
                linetype = as.factor(site_no),
                color = as.factor(site_no))) +
  ylab("Annual WY [acre-ft]") +
  theme_bw()


plotly::ggplotly(p)
```

```{r RndmStats, warning=FALSE, echo=FALSE}
#####
# calc and plot ecdfs
#####

# station IDs to calc ecdfs for
st_in <- unique(dt_dlyWY$site_no)
ecdfs <- data.table::rbindlist(
  lapply(
    st_in,
    function(x) {
      ecdf.fun(dt_dlyWY[site_no == x, X_00060_00003],
               ecdf.name = unique(dt_dlyWY[site_no == x, site_no]),
               plot.out = FALSE
               ) # close ecdf.fun
      tmp_cdf <- get(unique(dt_dlyWY[site_no == x, site_no]))
      tmp_cdf$name <- paste0(
        "st-", unique(dt_dlyWY[site_no == x, site_no])
        )
      tmp_cdf
    } # close function
  )
)

data_in <- ecdfs

p <- ggplot(data_in) +
  geom_line(aes(x = xs, y = Fx, color = name)) +
  theme_bw() +
  theme(legend.position = "none")

p
plotly::ggplotly(p)

st_in <- unique(dt_dlyWY$site_no)
skwn_out <- data.table(
  STAID = st_in,
  skwn = unlist(lapply(st_in, function(x){
      data_in <- dt_dlyWY[site_no == x, X_00060_00003]
      skewn.coef.fun(data_in, length(data_in), mean(data_in), sd(data_in))
      }
    )
  )
)
```

